---
title: "Introduction to cellkeyperturbation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to cellkeyperturbation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(cellkeyperturbation)
```
This vignette introduces the cellkeyperturbation package. The package contains the create_perturbed_table function which can be used to create a frequency table which has had a cell key perturbation technique applied to the counts according to rules supplied in a perturbation table.

The perturbation adds a small amount of noise to some cells in a table, changing their values. This adds uncertainty to small values to reduce the risk of disclosure. It protects against the risk of disclosure by differencing since it cannot be determined whether a difference between two similar tables represents a real person, or is caused by the perturbation.

Cell Key Perturbation is consistent and repeatable, so the same cells are always perturbed in the same way.

Data
-------

The create_perturbed_table function needs to be supplied with 

1. A data.table containing the data to be tabulated and perturbed. The data must contain a 'record key' variable and consist of one row per statistical unit (person, household, business or other) and one column per variable (age, sex, health status).

2. A data.table containing a p-table which provides the rules for the perturbation to be applied.

Two example data sets are supplied with this package:

micro: A small data set containing randomly generated data to showcase the cell key perturbation method.

ptable_10_5: A ptable containing the rules for applying a perturbation with a threshold of 10, and rounding to base 5.

Example
-------

```{r}
str(micro)

str(ptable_10_5)

perturbed_table <-create_perturbed_table(data = micro,
                                          record_key = "record_key",
                                          geog = c("var1"),
                                          tab_vars = c("var5","var8"),
                                          ptable = ptable_10_5)

str(perturbed_table)
```
